Draft: A missing guardrail---The minimal effect of election fraud on protest intensity in non-democracies

### 2022-11-30 Output/appendix2.Rmd
Edits to update v3 draft to MM instead of ECAV

models <- list()
models[['Model 1']] <- model.n.base.hurd
models[['Model 2']] <- model.n.base2.hurd
models[['Model 3']] <- model.n.griev.hurd
models[['Model 4']] <- model.n.griev2.hurd
models[['Model 5']] <- model.n.griev3.hurd

>>>>

models <- list()
models[['Model 1']] <- model.n.base.hurd.mm
models[['Model 2']] <- model.n.base2.hurd.mm
models[['Model 3']] <- model.n.griev.hurd.mm
models[['Model 4']] <- model.n.griev2.hurd.mm
models[['Model 5']] <- model.n.griev3.hurd.mm

### 2022-11-30 Output/manipulation_protest_draft_v3.Rmd
Edits to update v3 draft to MM instead of ECAV

Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, adapted from the recently-released Electoral Contention and Violence (ECAV) dataset [@Daxecker2019], this paper finds support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs. This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets. However, there is no statistically significant association between the level of election fraud and the number of protest events. This suggests a more complex relationship between election fraud and protest, undermining the argument that protest risk is a major deterrent to electoral malfeasance.

@@

```{r}
incl.vars <- nelda_massmob %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp,  median.participants.mm.pre, median.participants.mm.post)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))

datasummary_skim(incl.vars)
```

@@

In the same way, there is no statistically significant relationship between the level of fraud, the winning margin of victory, and the median size of post-election protest. This is true both for Model 6, where fraud is a standalone coefficient, and for the interaction term in Model 7. Figure \@ref(fig:vote-share-size-fig) shows no significant difference between the low-fraud and high-fraud conditions in any of the seven protest size categories, at any level of winning vote-share. Neither the models of protest size nor protest number support the strategic logic of protest advanced in some models, in which fraud sends a signal of weakness especially in the context of a weak ruling-party performance.

Turning to models of protest quantity presented in Table 3, there is no statistically significant relationship between fraud and protest in any of the models. Figure 4, using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share. As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory. Taking the results from Tables 2 and 3 together, there is support for Hypothesis 1 (an unconditional increase in the risk of any protest as fraud increases), as well as strong support for Hypothesis 2 (an increase in protest risk in fraudulent, low-margin elections) and Hypothesis 4 (no relationship between fraud and protest quantity). Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.

@@

```{r vote-share-size-fig, fig.cap="Fraud, winning vote-share, and median protest size", fig.height=7}
plot.combined.size <- cowplot::plot_grid(p.win.zero, p.win.low, p.win.high, ncol=1)
plot.combined.size
```

@@

```{r vote-share-fig, fig.cap="Fraud, winning vote-share, and number of protest events"}
p.mm.margin + theme_bw() + xlab("Winning vote-share") + ylab("Predicted number") + labs(title = "Predicted number of post-election protest", color = "Election fraud")  
```

>>>>

Using cross-national data on 647 elections from 1990 to 2012 where incumbents claimed victory, combining data from NELDA [@hyde_2012_marinov], V-Dem [@vdem11.1], and the Mass Mobilization dataset [@massmob], this paper finds support for the hypothesis that more fraudulent elections are associated with a higher risk that any post-election protest event occurs. This is in keeping with the formal-modeling literature, which predicts that overreach in the form of election fraud indicates government weakness, making opposition groups confident of winning a conflict in the streets. However, there is no statistically significant association between the level of election fraud and the number of protest events. This suggests a more complex relationship between election fraud and protest, undermining the argument that protest risk is a major deterrent to electoral malfeasance.

@@

```{r, include=F}
incl.vars <- nelda_massmob %>% dplyr::select(jucon.1lag, legcon.1lag, altinf.1lag, gdpgro.1lag, osorg.1lag, pubseccorrup.lag, econ.crisis.nelda.1lag, natlpartyorg.lag,  pubseccorrup.lag,  pubseccorrup.lag, v2elirreg.inv, presidential, inc.termlimit.num, physinteg.1lag, urban.pct, rentierstate, taxstate, winner.share.upd, harass_opp,  median.participants.mm.pre, median.participants.mm.post)
#incl.vars <- incl.vars %>% mutate(v2elintmon = as.numeric(v2elintmon))
```

```{r}
datasummary_skim(incl.vars)
```

@@

In the same way, there is no statistically significant relationship between the level of fraud, the winning margin of victory, and the median size of post-election protest. This is true both for Model 6, where fraud is a standalone coefficient, and for the interaction term in Model 7. Figure \@ref(fig:vote-share-size-fig) shows no significant difference between the low-fraud and high-fraud conditions in any of the seven protest size categories, at any level of winning vote-share. That is, the severity of fraud does not appear associated with changes in protest size. As a whole, neither the models of protest size nor protest number support the strategic logic of protest advanced in some models, in which fraud sends a signal of weakness especially in the context of a weak ruling-party performance.

Turning to models presented in Table 3, there is no statistically significant relationship between fraud and the number of protest events in any of the models. Figure \@ref(vote-share-fig), using results from Model 9, presents the marginal effects for fraud conditional on the incumbent's vote-share. As the figure shows, there is no statistically significant difference in the expected number of protest following a relatively clean election compared to a relatively fraudulent one, for any margin of incumbent victory.

Taking the results from Tables 2 and 3 together, Hypothesis 1\textemdash on the unconditional relationship between increasingly severe fraud and the likelihood of protest\textemdash is not supported. Alone, fraud severity appears to have no relationship with either the size or quantity of protest. Similarly, there is no support for Hypothesis 2, which predicts an increase in protest risk in fraudulent, low-margin elections.  and Hypothesis 4 (no relationship between fraud and protest quantity). Hypothesis 3, on the connection between fraud, grievance, and protest, is not supported.

@@

```{r vote-share-size-fig, fig.cap="Fraud, winning vote-share, and median protest size", fig.height=7, echo=FALSE, message=FALSE}
plot.combined.size <- cowplot::plot_grid(p.win.zero, p.win.low, p.win.high, ncol=1)
plot.combined.size
```

@@

```{r vote-share-fig, fig.cap="Fraud, winning vote-share, and number of protest events",  echo=FALSE, message=FALSE}
p.mm.margin + theme_bw() + xlab("Winning vote-share") + ylab("Predicted number") + labs(title = "Predicted number of post-election protest", color = "Election fraud")  
```
